Sperm modifiers of female drive.

### 2013-09-11 writeup_and_notes/male_dependent_drive.tex
added invasion figure file
added discussion of homozyg. cost invasion to paper

YB: I think we should do our best to be very short and sweet here. There are many opportunities to move off to cool math and cool biological scenarios. 
We should fight all of them and focus on mission critical results. All others can go in supp or appease reviewers. Lets talk about why the critical results are. To me its (1) male enhancement of female drive cannot maintain a stable polymorphisms, and (2) An allele in sperm can evolve to suppress its drive in females.

Rough order of rest of text
\subsection*{Lay out Model} 
We imagine that drive acts in such a way that an allele promote itself into a fraction of d
 eggs ($d> 1/2$, when it is present in a heterozygote, forcing the other allele to be present in $1-d$ eggs. 
In mammals, fertilization takes place at MII, so we imagine this drive must be taking place at MII in order
for a sperm to have any influence. For drive to take place at MII there has to be an uneven number of
 crossovers between the centromere and the drive locus, such that realistically $d$ is bounded to be $<XX$. 
In other systems where fertilization occurs during MI, sperm could influence either drive at MI or MII, 
and drivers at MI can have a $d=1$ if they occur in tight linkage with the centromere.

Cartoon figure of sperms meeting eggs with different genotypes?

\begin{figure}
	\rotatebox{270}{\includegraphics[width = 0.8
          \textwidth]{Figures/sperm_egg_cartoon1.ps}}
\caption{transmission probabilities for alleles through female
  meiosis depend on sperm genotype. 2 allele models}  
\label{Eggsperm_2_allele_cartoon}
\end{figure}

\begin{figure}
	\rotatebox{270}{\includegraphics[width = 0.8
          \textwidth]{Figures/sperm_egg_cartoon1.ps}}
\caption{transmission probabilities for alleles through female
  meiosis depend on sperm genotype. 2 allele models}  
\label{Eggsperm_2_allele_cartoon}
\end{figure}

@@

\subsection*{ A) Invasion by single self promoting allele}
show that such alleles:
\begin{enumerate}
\item can't be balanced, \\
\item and homozygous problems are tested out at low freq.  \\
\item Any heterozygous problems, leads to a bistable allele\\
\item If these alleles take off they speed through to fixation\\
\item If allele has any drive ability in absence of sperm effect that is what allows it to enter the population
and sperm effect isn't a further cause of conflict. What if anything do we mean by this?\\
\item PERHAPS HERE WE INTRODUCE A SELF-RESTRAINING ALLELE
\end{enumerate}

>>>>

\yb{YB: To me its (1) male enhancement of female drive cannot maintain a stable polymorphisms, and (2) An allele in sperm can evolve to suppress its drive in females.}

Rough order of rest of text
\subsection*{A single locus model} 
The standard model of female meiotic drive is two allele model, where the
allele $1$ is transmitted $d$ fraction of the time through female
meiosis ($d > \half$), regardless of the genotype of the sperm. 
In mammals, fertilization takes place at MII, so we imagine this drive must be taking place at MII in order
for a sperm to have any influence. For drive to take place at MII there has to be an uneven number of
 crossovers between the centromere and the drive locus, such that realistically $d$ is bounded to be $<0.75$?. 
In other systems where fertilization occurs during MI, sperm could influence either drive at MI or MII, 
and drivers at MI can have a $d=1$ if they occur in tight linkage with the centromere.

We can model a simple sperm dependent meiotic drive system, by
modifying this model so that the allele only drives during female
meiosis if it fertilized by a sperm carrying that allele
 (see
Figure \ref{Eggsperm_2_allele_cartoon}). Under what conditions can
this self propograting driver allele spread, and could it drive the
evolution of supressors?

We first consider the situation where the driving allele has no
fitness consequences in heterozygotes, only in homozygotes (whose
relative fitness is $1-s$). A standard meiotic drive allele can always
invade the population in this situation, as it is initially only
present mostly in heterozygotes and rarely suffers the consequences of
the problems in homozygotes. However, if the cost in homozygotes is
strong enough ($s>XXd$), then the allele is prevented from reaching
fixation and instead is maintained as a balanced polymorphism in the
population \citep{Prout1973}. Such balanced polymorphisms can then
drive the evolution of supressors of drive \citep{XX}, and most of our
examples of drive systems are of this form.

Under a recessive fitness cost model our sperm dependent driver
actually has far more restrictive conditions to invade the population,
than a standard meiotic driver (see Figure \ref{Invasion_space}).
That is because our allele which drives only when it is present in the
fertilizing sperm is guarenteed to form a homzygote. Thus, 
to even spread in the population when it is rare, our sperm-dependent
driver has to overcome the cost it suffers in the homozygote state.

When conditions are suitable for such alleles to spread into the
population, they spread slowly through low frequencies as they rarely
drive as they are unlikely to be fertilized by a sperm with the right
allele. Once they become established they spread rapidly due to their
positive frequency dependent behavior (Inset in Figure \ref{Invasion_space}). 
Unlike standard meiotic drive systems 
these alleles can not become balanced in the population  by homozygote
cost, as they have already over come this cost just to enter the population.

HETERO

\begin{figure}
	\rotatebox{270}{\includegraphics[width = 0.8
          \textwidth]{Figures/sperm_egg_cartoon1.ps}}
\caption{transmission probabilities for alleles through female
  meiosis depend on sperm genotype. 2 allele model}  
\label{Eggsperm_2_allele_cartoon}
\end{figure}

\begin{figure}
\includegraphics[width = 0.8 \textwidth]{Figures/invasion_space_recessive_driver.eps}
\caption{Invasion analysis. Likely merge with Figure 1.} \label{Invasion_space}
\end{figure}

@@

\subsection*{ A) Invasion by single self promoting allele}

show that such alleles:
\begin{enumerate}
\item can't be balanced, \\
\item and homozygous problems are tested out at low freq.  \\
\item Any heterozygous problems, leads to a bistable allele\\
\item If these alleles take off they speed through to fixation\\
\item If allele has any drive ability in absence of sperm effect that is what allows it to enter the population
and sperm effect isn't a further cause of conflict. What if anything do we mean by this?\\
\item PERHAPS HERE WE INTRODUCE A SELF-RESTRAINING ALLELE
\end{enumerate}
